---
title: "Getting Started with cBioPortalData"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to use the packaged-data and API-data functions to retrieve data
  from cBioPortal using the cBioPortalData Bioconductor package.
---

```{r setup, include=FALSE}
library(learnr)
library(cBioPortalData)
knitr::opts_chunk$set(echo = FALSE)
```

## cBioPortalData Tutorial

In this tutorial, you will learn when and how to use

* `cBioDataPack` for pre-packaged data
* `cBioPortalData` for API-query data

**Note**. See the source code of the interactive exercises at
https://github.com/waldronlab/MultiAssayWorkshop

### What data is available?

First we need to know if we are interested in any of the data that is offered
by the cBio Genomics Portal.

### getStudies()

We can use the `getStudies` function to get a list of available studies from
the API. Enter `cbio` as the input to `getStudies()`. We also want the build
reporting enabled with `buildReport = TRUE`.

```{r getstudies, exercise=TRUE, exercise.lines = 3}
cbio <- cBioPortal()
getStudies(buildReport = TRUE)
```

```{r getstudies-solution}
getStudies(api = cbio, buildReport = TRUE)
```

###

We are interested in the first and second to last columns in the table, namely
`studyId` and `pack_build` / `api_build`.


## Pre-packaged Data

### cBioDataPack

Now that we have an idea of how to see what studies are documented and
available for download, we can bring one in. Let's use the `cBioDataPack`
function to download the Acute Lymphocytic Lukemia from St. Jude dataset with
the study identifier: `"all_stjude_2016"`

```{r stjude, exercise=TRUE, exercise.eval=FALSE}
"all_stjude_2016"
```

```{r stjude-solution}
cBioDataPack(cancer_study_id = "all_stjude_2016")
```

###

Great! We just loaded our first packaged dataset. Note that not all packages
currently build as `MultiAssayExperiment` objects. Refer to the `pack_build`
column in `getStudies()`, with the `buildReport = TRUE` enabled, to see which
study identifiers are working as `MultiAssayExperiment` represenations!

## API Data

### cBioPortalData

We can also request more granular data from `cBioPortal` data service.
First we need an API object for so that R knows how and where to send queries.
We create one using the `cBioPortal` function. Use it and assign it to an
object called `cbio`. You can have a look at the API object by calling it
directly (type and submit the name to the console).

```{r cbioportal, exercise = TRUE, exercise.lines = 2}

```

```{r cbioportal-solution}
cbio <- cBioPortal()
```

### Sending a query

First let's look at what pieces of information we need to write a successful
`cBioPortalData` function call.

We need to know:
* `studyId` - the study identifier from `getStudies`
* `genePanelId` or `genes` - what features we are interested in
* `by` - whether we want these features as Entrez IDs or Hugo Symbols.
* `molecularProfileIds` - specific assays that make up a study

```{r,echo=FALSE}
cbio <- cBioPortal()
```

### genePanels

Use the `genePanels` function to see what gene panels are available (if
desired).

```{r genepanel, exercise = TRUE, exercise.eval = FALSE}
genePanels(cbio)
```

### genes

We can have a look at a table of genes with the `geneTable` function.

```{r genetable, exercise = TRUE, exercise.eval = FALSE}
geneTable(cbio)
```

Notice that the output changes by page number and based on the `pageSize`
parameter.

If you have a specific set of genes in mind, those can be used as input into
the main `cBioPortalData` function. We will see how in a few moments.

### molecularProfileIds

Use the `molecularProfiles` function to see all the assays available for a
particular study. In this exercise, let's look at the `molecularProfiles`
table available for Adrenocortical Carcinoma or ACC with the `acc_tcga` study
identifier. We are particularly interested in the `molecularProfileId` column
in the output table.

```{r molprofs, exercise = TRUE, exercise.eval = FALSE}
molecularProfiles()
```

```{r molprofs-solution}
molecularProfiles(cbio, studyId = "acc_tcga")
```


### Bringing it together

Now that we have all the pieces, we can construct the `cBioPortalData` function
call based on the examples provided. We want to look at the `acc_tcga` `studyId`
along with a couple of `molecularProfileIds` say `'acc_tcga_rppa'` and
`'acc_tcga_mutations'`. For this example, let's use `'IMPACT341'` as the
`genePanelId` of interest. Let's put all of these in the function.

```{r querycbio, exercise = TRUE, exercise.eval = FALSE}
cBioPortalData(
    cbio,
    studyId = ,
    genePanelId = , # OR genes =
    molecularProfileIds = c(),
    by = "entrezGeneId"
)
```

```{r querycbio-solution}
cBioPortalData(
    cbio,
    studyId = "acc_tcga",
    genePanelId = "IMPACT341",
    molecularProfileIds = c("acc_tcga_rppa", "acc_tcga_mutations"),
    by = "entrezGeneId"
)

## example with genes argument
cBioPortalData(
    cbio,
    studyId = "acc_tcga",
    genes = c(25, 84142, 207, 208, 10000, 238, 242, 139285, 324, 367),
    molecularProfileIds = c("acc_tcga_rppa", "acc_tcga_mutations"),
    by = "entrezGeneId"
)
```

###

Amazing! We now have a working dataset obtained from the cBioPortal API.

## Trivia!

Let's understand the difference between the two primary function in the
`cBioPortalData` package:

```{r cbio-quiz}
quiz(
  question("What function provides pre-packaged cBioPortal study data in its entirety?",
      answer("data",
          message = "Oops! This function lets you see what data are available either in R or in a package."),
      answer("cBioDataPack", correct = TRUE, message = "Nice one!"),
      answer("cBioPortal",
          message = "Oh no! This function only provides the API representation of the service."),
      answer("load", message = "Yikes! The load function lets users load serialized binary data as written by the save function."),
      allow_retry = TRUE
  ),
  question("What function can you use to obtain the cBioPortal data as served by the API service?",
      answer("cBioPortal",
          message = "Oh no! This function only provides the API representation of the service."),
      answer("cBioPortalData", correct = TRUE, message = "You're on fire!"),
      answer("molecularProfiles",
          message = "D'oh! This function allows you to see what assays are available for a particular study ID."),
      answer("cBioDataPack",
          message = "Close! This function provides the pre-packaged cBioPortal data."),
      allow_retry = TRUE
  ),
  question("What component(s) do we need to use the cBioPortalData function?",
      answer("studyId"),
      answer("genes or genePanelId"),
      answer("molecularProfileIds"),
      answer("All of the above", correct = TRUE, message = "Great job!")
  )
)
```

